\input{settings} % add packages, settings, and declarations in settings.tex
\usepackage{url}
\usepackage{amsmath,amssymb,amsthm}


\newcommand{\Prob}[1]{{{\bf{Pr}}\left[{#1}\right]}}
\newcommand{\Mean}[1]{{\mathbb E}\left[{#1}\right]}
\newcommand{\Var}[1]{{\mathbb Var}\left[{#1}\right]}


\begin{document}

\lhead{Prof. C.E. Tsourakakis} 
\rhead{CS365 Spring '22 \\ Foundations of Data Science \\ Assignment 10} 
\cfoot{\thepage\ of \pageref{LastPage}}


\section*{Instructions}
\framebox{%
	\begin{minipage}{0.9\linewidth}
		\begin{itemize}
			\item The homework is due on \underline{{\bf Friday 4/29 at 5pm ET}}.  
			\item No extension will be provided, unless for serious documented reasons.
			\item {\bf Start early!}
			\item Study the material taught in class, and feel free to do so in small groups, but the solutions should be a product of your own work. 
			\item This  is not a multiple choice homework;   reasoning, and mathematical proofs are required before giving your final answer.
			\item  $tr$ denotes  the trace of a matrix. 
		\end{itemize}
\end{minipage}}

\section{Derivatives [55 points]} 

Compute the derivative $\frac{df}{dx}$   for the following functions. It will be helpful to  identify $n,m$ where $f:\mathbb{R}^n \to \mathbb{R}^m$, and the dimensions of the derivative first. 

\begin{enumerate}
	\item [(a)] [5pts]  $f(x) = \frac{1}{1+e^{-x}}, x \in \mathbb{R}$
	
	\vspace{2mm} 

	\item[(b)]  [5 pts]  $f(x)= \exp \Big(-\frac{1}{2\sigma^2} (x-\mu)^2 \Big), x \in \mathbb{R}$
	
 
	 

	\item[(c)]  [5 pts]   $f(x)= \sin (x_1) \cos (x_2), x \in \mathbb{R}^2$.
	
	
	
	\item[(d)]  [5 pts]    $f(x) = xx^T, x \in \mathbb{R}^n$.

	\item[(e)]  [5 pts]    $f(x) = \sin( \log (x^T x)), x \in \mathbb{R}^n$.  
 
 	\item[(f)]  [10 pts]    $f(z)=\log(1+z)$ where $z=x^Tx, x \in \mathbb{R}^n$

 	\item[(g)]  [10 pts]    $f(x)=tr(xx^T+\sigma^2 I), x \in \mathbb{R}^n$.  
	
	\item[(h)] [10 pts] $f(x)=x^TAx$ where $x \in \mathbb{R}^n, A \in \mathbb{R}^{n \times n}$.
\end{enumerate}   

 

\section{Matrix derivative [25 points]}
 
 Let $A \in \mathbb{R}^{D \times E}, X \in  \mathbb{R}^{E \times F}, B \in  \mathbb{R}^{F \times D}$. Compute $\frac{dg}{dX}$ where  $g(X) = tr(AXB).$
 
 
\section{SVD Practice [20 points]}

Find the  SVD of the following matrices without the use of computing devices/software.


\begin{enumerate}
    \item[(a)] [5 points]  $A= [1 , 1]$.
    \item[(b)]{[5 points] $ B=
    	 \begin{bmatrix}
        1 & 3 \\
        2 & 6 \\
         
        \end{bmatrix}. $
    }
    \item[(c)]{[5 points] $C= 
         \begin{bmatrix}
        1 & 2 \\
        3 & 6 \\
         
        \end{bmatrix}.$
    }
    \item[(d)]{[5 points] 
        $D= \begin{bmatrix}
        1 & 2 & 3\\
        2 & 4 & 6\\
        \end{bmatrix}.$
    }
\end{enumerate}

\end{document}
