\input{settings} % add packages, settings, and declarations in settings.tex
\usepackage{url}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}

\newcommand{\Prob}[1]{{{\bf{Pr}}\left[{#1}\right]}}
\newcommand{\Mean}[1]{{\mathbb E}\left[{#1}\right]}
\newcommand{\Var}[1]{{\mathbb Var}\left[{#1}\right]}


\begin{document}

\lhead{Prof. C.E. Tsourakakis} 
\rhead{CS365 Spring '22 \\ Foundations of Data Science \\ Assignment 5} 
\cfoot{\thepage\ of \pageref{LastPage}}


\section*{Instructions}
\framebox{%
	\begin{minipage}{0.9\linewidth}
		\begin{itemize}
			\item The homework is due on \underline{{\bf Friday 3/18 at 5pm ET}}.  
			\item There are 3 problems in total.  
			\item No extension will be provided, unless for serious documented reasons.
			\item {\bf Start early!}
			\item Study the material taught in class, and feel free to do so in small groups, but the solutions should be a product of your own work. 
			\item This  is not a multiple choice homework;   reasoning, and mathematical proofs are required before giving your final answer.
		\end{itemize}
\end{minipage}}

\section{Reservoir Sampling [30 Points]} 

Design an algorithm that samples  $k \geq 1$ elements uniformly at random  from an insert-only stream, whose length is uknown. Present the pseudocode and the proof of the algorithm's correctness.

\section{Variance of Morris Counter [35 points]} 

Prove equation $\text{Var}(Z)=\frac{m(m-1)}{2}$ on \href{https://tinyurl.com/2p84tftt}{slide 27}.

\section{A useful technique  [35 points]} 

Prove the claim on \href{https://tinyurl.com/putc3vw2}{slide 16}.  Be specific about the values of the constants $C_1, C_2$ you use in your proof, where $t=C_1 \log \frac{1}{\delta}$, $k=C_2 \frac{\text{Var}[X]}{\epsilon^2 \mathbb{E}[X]^2}$.
 
\end{document}
